---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(udmpullr)
library(magrittr)
```

# udmpullr

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of udmpullr is to provide convenient functions to access data from the [Understanding Data and Markets (UDM) API](https://commoditymarkets.com/).

## Installation

You can install the development version of udmpullr from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tmhanon/udmpullr")
```

or

``` r
# install.package("devtools")
devtools::install_github("tmhanon/udmpullr")
```

## Basic Workflow

### Setting Credentials

```{r include = FALSE}
has_auth <- httr2::secret_has_key("UDMPULLR_KEY") && !identical(Sys.getenv("GITHUB_WORKFLOW"), "R_CMD-check")

my_UDM_username <- httr2::secret_decrypt("rZDs3tqcTToVAuIVmJwHj-G7dFk8u2odlrmYpF0Uh1r-oc2_zQY", "UDMPULLR_KEY")
my_UDM_password <- httr2::secret_decrypt("1oLFr_3eTNUd4-G_w8QrIX7WK-mKyD4q--KIyyO3", "UDMPULLR_KEY")
```

In order to access the UDM API you must have an account registered with UDM. Visit <https://commoditymarkets.com/> for more information. Once your account is registered, you will use your username (typically an email address) and password to authorize access to the UDM API. There are multiple options to set your credentials to allow `udmpullr` to pull from the UDM API on your behalf:

1.  Use `set_udm_credentials()` to input your credentials (username and password). R will save your credentials and `udmpullr` will be able to access them within the same session. However, next time you open R you will have to re-enter your credentials. `set_udm_credentials()` takes two arguments, `UDM_API_USERNAME` and `UDM_API_PASSWORD`. Alternatively, you can leave either or both arguments blank and R will prompt you to enter your credentials in popup windows (using the [`askpass`](https://github.com/r-lib/askpass) package).

```{r}
set_udm_credentials(UDM_API_USERNAME = my_UDM_username,
                    UDM_API_PASSWORD = my_UDM_password)

# set_udm_credentials()
# Running without arguments will prompt you to enter your credentials in a popup window.
```

2.  If you use the UDM API regularly, a better option is to use environmental variables to permanently set your credentials. Open or create a `.Renviron` file using `usethis::edit_r_environ()` and add `UDM_API_USERNAME="<your UDM email/username>"` and `UDM_API_PASSWORD="<your UDM password>"` on separate lines. Save the file (ensuring there is a blank line at the end of the file) and restart R. Your 'UDM' API credentials will now be available to `udmpullr()` across R sessions.

After you have set your UDM API credentials using either of these options, they can be retrieved using the functions `get_udm_username` and `get_udm_password`. The main `udmpullr()` function uses these retrieval functions as default arguments so access to the UDM API is authorized correctly.

### Accessing Data 

The primary function to access data from the UDM API is `udmpullr()`. Suppose you want to access the daily dairy product prices from the CME spot market. Within the UDM API, these data are contained within a table named "UDM_Report_CME_Dairy_Spot_Market_Overview_Daily". This can be passed to `udmpullr()` as the first argument, `table_name`:

```{r}
udmpullr(table_name = "UDM_Report_CME_Dairy_Spot_Market_Overview_Daily")
```

As shown, `udmpullr()` will return the entire table, with data back to 1998. In most cases this is likely overkill, so `udmpullr()` also allows you to specify a start date and end date using the `start_date` and `end_date` variables. This filters the data before it is pulled, which should also speed up the response time. The following example will pull data for the last two weeks (note that using `Sys.Date()` as the `end_date` argument is not strictly necessary).

```{r}
udmpullr(table_name = "UDM_Report_CME_Dairy_Spot_Market_Overview_Daily",
         start_date = Sys.Date() - lubridate::days(14),
         end_date = Sys.Date())
```

The UDM API has a limit of 10,000 rows of data in any single response. This is an issue for certain tables, such as the daily futures panel, which have far more than 10,000 rows of data. Therfore, it is necessary to filter the data to some extent before pulling from the UDM API. The `start_date` and `end_date` arguments are one option for doing so, but `udmpullr()` also allows specific products to be specified (for tables where that is applicable). For example, the daily futures panel includes a variable called `Product_Name`, and `udmpullr()` allows a specific product to be specified via the `product_name` argument:

```{r}
udmpullr("UDM_Report_Futures_Prices_Daily_Panel",
         product_name = "CME Class III Milk Futures",
         start_date = Sys.Date() - lubridate::days(7))
```

The `product_name` argument can also easily be iterated over using the ``purrr::map()` function, allowing for several products to be pulled simultaneously.

```{r}
list("CME Class III Milk Futures",
     "CME Class IV Milk Futures",
     "CME Cash-Settled Butter Futures",
     "CME Non-Fat Dry Milk Futures",
     "CME Dry Whey Futures",
     "CME Cash-Settled Cheese Futures") %>%
  purrr::map(\(x) udmpullr("UDM_Report_Futures_Prices_Daily_Panel",
                           product_name = x,
                           start_date = Sys.Date() - lubridate::days(7),
                           end_date = Sys.Date())) %>%
  dplyr::bind_rows()
```

